---
title: "Incytr_tutorial_5XAD_202412"
author: "Changhan"
date: "2024-12-09"
output: html_document
---

################################################################################
Note: This code is unpublic and intended for use within the [Incytr/Signaling networks discovery paper] project only.
Please refrain from sharing it outside the project team or using it for any other projects without prior approval from the author.
################################################################################

# Preparation

libraries
```{r}
library(ggplot2)
library(plotrix)
library(parallel)
library(doParallel)
```


RNA-seq data
```{r}
# path to the folder
setwd("C:/Users/chang/Documents/Research_2023/CST_project/CST_Project/CST_scRNAseq_data")
Data.input <- readRDS("Az_8clusteres.RDS") 
```

Marker genes
```{r}
setwd("C:/Users/chang/Documents/Research_2023/CST_project/CST_Project/DG_analysis")
Allmarkers = read.csv("Allmarkers_18groups.csv")
```

Multimodel data
```{r}
setwd("C:/Users/chang/Documents/Research_2023/CST_project/CST_Project/CST_processed_data")

pr_5X = read.csv("processed_pr_5X_v2.csv")
pr_WT = read.csv("processed_pr_WT_v2.csv")
ps_5X = read.csv("processed_ps_5X_v2.csv")
ps_WT = read.csv("processed_ps_WT_v2.csv")
py_5X = read.csv("processed_py_5X_v2.csv")
py_WT = read.csv("processed_py_WT_v2.csv")

kldata = read.csv("kldata.csv")
```

Incytr database
```{r}
setwd("C:/Users/chang/Documents/Research_2023/CST_project/CST_Project/Database")
# load the exFINDER database (can also download the updated database from the exFINDER github page)
load("DB_Layer1_mouse_filtered.rda")
load("DB_Layer2_mouse_filtered.rda")
load("DB_Layer3_mouse_filtered.rda")

DB.M <- list()
DB.M[[1]] <- DB_Layer1_mouse_filtered
DB.M[[2]] <- DB_Layer2_mouse_filtered
DB.M[[3]] <- DB_Layer3_mouse_filtered
```

# Testing

The following chunk can be used for testing (only analyze the signaling pathways between two selected clusters).
Permutation test = 10 
```{r}
start.time <- Sys.time()

setwd("C:/Users/chang/Documents/Research_2023/CST_project/CST_Project/output")

cell.groups = levels(Data.input$Type)

j = 3
i = 1
# k =13

Sender.group = cell.groups[i]
Receiver.group = cell.groups[j]
# set the Scenario
Scenario = 4
# input genes are DEG
DG.Sender <- Allmarkers$gene[Allmarkers$cluster %in% c( paste0(Sender.group, "_5X"),paste0(Sender.group, "_WT") )]
DG.Receiver <- Allmarkers$gene[Allmarkers$cluster %in% c( paste0(Receiver.group, "_5X"),paste0(Receiver.group, "_WT") )]

# Creat object
Incytr <- create_Incytr(object = Data.input@assays$RNA@data,
                        meta = Data.input@meta.data,
                        sender = Sender.group,
                        receiver = Receiver.group,
                        group.by = "Type",
                        condition = c("5X", "WT"),
                        assay = NULL,
                        do.sparse = T)

# Infer paths
Incytr <- pathway_inference(Incytr,
                            DB = DB.M,
                            gene.use_Sender = DG.Sender,
                            gene.use_Receiver = DG.Receiver,
                            ligand = NULL,
                            receptor = NULL,
                            em = NULL,
                            target = NULL)

# Calculate the average expression level
Incytr <- Expr_bygroup(Incytr,
                       mean_method = NULL)

# Predict signaling prob. (from transcriptomics data)
# "cutoff_SigWeight = 0.10": keep the pathways that SigWeight>=0.10 in at least one condition
try.function <- try(     Incytr <- Cal_SigProb(Incytr, K = 0.5, N = 2, cutoff_SigProb = 0.10,
                                               correction = 0.001, q = NULL) )
if( "try-error" %in% class(try.function) ){
  if( isTRUE(grepl(text, try.function[1])) ){
    next
  }
}

# Integrate the multiomics data (depend on the Scenario)
if(Scenario == 1){ # no multimodel data
  Incytr <- Integr_multiomics(Incytr, 
                              pr.data_condition1 = NULL,
                              pr.data_condition2 = NULL,
                              pr.correction = NULL,
                              pr.q = NULL,
                              ps.data_condition1 = NULL,
                              ps.data_condition2 = NULL,
                              ps.correction = NULL,
                              ps.q = NULL,
                              py.data_condition1 = NULL,
                              py.data_condition2 = NULL,
                              py.correction = NULL,
                              py.q = NULL)
}else if(Scenario == 2){ # proteomics data
  Incytr <- Integr_multiomics(Incytr, 
                              pr.data_condition1 = pr_5X,
                              pr.data_condition2 = pr_WT,
                              pr.correction = NULL,
                              pr.q = NULL,
                              ps.data_condition1 = NULL,
                              ps.data_condition2 = NULL,
                              ps.correction = NULL,
                              ps.q = NULL,
                              py.data_condition1 = NULL,
                              py.data_condition2 = NULL,
                              py.correction = NULL,
                              py.q = NULL)
}else if(Scenario == 3){ # ps data
  Incytr <- Integr_multiomics(Incytr, 
                              pr.data_condition1 = pr_5X,
                              pr.data_condition2 = pr_WT,
                              pr.correction = NULL,
                              pr.q = NULL,
                              ps.data_condition1 = ps_5X,
                              ps.data_condition2 = ps_WT,
                              ps.correction = NULL,
                              ps.q = NULL,
                              py.data_condition1 = NULL,
                              py.data_condition2 = NULL,
                              py.correction = NULL,
                              py.q = NULL)
}else if(Scenario == 4){# py data
  Incytr <- Integr_multiomics(Incytr, 
                              pr.data_condition1 = pr_5X,
                              pr.data_condition2 = pr_WT,
                              pr.correction = NULL,
                              pr.q = NULL,
                              ps.data_condition1 = ps_5X,
                              ps.data_condition2 = ps_WT,
                              ps.correction = NULL,
                              ps.q = NULL,
                              py.data_condition1 = py_5X,
                              py.data_condition2 = py_WT,
                              py.correction = NULL,
                              py.q = NULL)
}

# Calculate the final score
# "cutoff_RNA_score": keep the pathways that "abs(RNA_score)>=cutoff_RNA_score"
try.function <- try(  Incytr <- Pathway_evaluation(Incytr, score.weight = NULL, k_logi = 2, style = NULL, abs.value = NULL,
                                                   cutoff_TPRS = NULL) )
if( "try-error" %in% class(try.function) ){
  if( isTRUE(grepl(text, try.function[1])) ){
    next
  }
}

Incytr <- Integr_kinasedata(Incytr,
                            kldata = kldata,
                            mean_method = NULL,
                            cell_group = levels(Incytr@meta$Type),
                            fold_threshold = 10)

# Combine the Kinase score with previous calculated "final score"
# "cutoff_final_score = 0.2": keep the pathways that "abs(final_score)>=0.2"
try.function <- try( Incytr <- Cal_PRS(Incytr, KPRS.weight = 0.5,
                                       cutoff_PRS = 0.2) )
if( "try-error" %in% class(try.function) ){
  if( isTRUE(grepl(text, try.function[1])) ){
    next
  }
}

# "cutoff_p_value = 0.1": keep the pathways that "p_value<=0.1" in at least one condition
try.function <- try( Incytr <- Permutation_test(Incytr,
                                                nboot = 10,
                                                seed.use = 1L,
                                                mean_method = NULL,
                                                cutoff_p_value = 0.1) )
if( "try-error" %in% class(try.function) ){
  if( isTRUE(grepl(text, try.function[1])) ){
    next
  }
}

# infer SrK (a seperate data.frame which includes both SiK and SrK, saved at Incytr@kl.explore)
Incytr <- Kinase_exploration(Incytr,
                             kldata = kldata,
                             mean_method = NULL, 
                             cell_group = levels(Incytr@meta$Type),
                             fold_threshold = 10)

# Export the results (does not include the SrK analysis results)
output <- Export_results(Incytr,
                         indicator = TRUE)

end.time <- Sys.time()
time.1 <- end.time - start.time
time.1
```

# Identify highly expressed genes (>75th percentile)
```{r}
setwd("C:/Users/chang/Documents/Research_2023/CST_project/CST_Project/CST_processed_data")

cell.groups = levels(Data.input$Type)

HEG.df = data.frame(gene_symbol = as.character(), ave.exp = as.numeric(), cluster = as.character(), condition = as.character())

for (k in 1:length(conditions)) {
  for (i in 1:length(cell.groups) ) {
    # seperate the data by the conditions
    HGData.input = subset(x = Data.input, subset = condition == conditions[k])
    
    # for select cluster, find genes that: average expression level >= cutoff_exp
    HE <- Find_highexp_gene(object = HGData.input@assays$RNA@data,
                            meta = HGData.input@meta.data,
                            group.by = "Type",
                            group.select = cell.groups[i],
                            geneuse = NULL,
                            cutoff_percentile = 0.75,
                            mean_method = NULL)
    
    df.new = data.frame(gene_symbol = HE[ , 1], ave.exp = HE[ , 2], cluster = cell.groups[i], condition = conditions[k])
    HEG.df = rbind(HEG.df, df.new)
  }
}

write.csv(HEG.df, file="HE_gene_075_121024.csv")
```


# Identify differentially (proteomics) expressed genes (fold change>1)
```{r}
setwd("C:/Users/chang/Documents/Research_2023/CST_project/CST_Project/CST_processed_data")

pr.gene_output = proteomics_gene(pr.data_condition1 = pr_5X, 
                                 pr.data_condition2 = pr_WT,
                                 cell_group =  levels(Data.input$Type),
                                 style = "aFC",
                                 cutoff = 1,
                                 pr.correction = 0.00001,
                                 pr.q = NULL)

write.csv(pr.gene_output, file="proteomics_gene_121024.csv")
```


Load HEG & proteomics genes
```{r}
setwd("C:/Users/chang/Documents/Research_2023/CST_project/CST_Project/CST_processed_data")

HEG.df = read.csv("HE_gene_075_121024.csv")
pr.gene_output = read.csv("proteomics_gene_121024.csv")
```

################################################################################

# Incytr analysis 1: input genes = DEG + HEG


```{r}
text = "No signaling pathway inferred using current filter, the filter has been ignored."
```


Vareables to store pairwise-analysis results
```{r}
Pairwise_pathway_table = list()
Pairwise_kinase_table = list()
```


```{r, include=FALSE}
start.time <- Sys.time()

setwd("C:/Users/chang/Documents/Research_2023/CST_project/CST_Project/output_DEG_HEG")

cell.groups = levels(Data.input$Type)

k = 1

for (i in 1:length(cell.groups) ) {
  for (j in 1:length(cell.groups) ) {
    # select the sender & receiver groups
    Sender.group = cell.groups[i]
    Receiver.group = cell.groups[j]
    # set the Scenario
    Scenario = 4
    # input genes are DEG
    DG.Sender <- Allmarkers$gene[Allmarkers$cluster %in% c( paste0(Sender.group, "_5X"),paste0(Sender.group, "_WT") )]
    DG.Receiver <- Allmarkers$gene[Allmarkers$cluster %in% c( paste0(Receiver.group, "_5X"),paste0(Receiver.group, "_WT") )]
    
    ############################################################################
    # Add highly expressed genes
    HG.Sender <- HEG.df$gene_symbol[HEG.df$cluster == Sender.group]
    HG.Receiver <- HEG.df$gene_symbol[HEG.df$cluster == Receiver.group]
    
    DG.Sender <- unique(c(DG.Sender, HG.Sender))
    DG.Receiver <- unique(c(DG.Receiver, HG.Receiver))
    
    # Create object
    Incytr <- create_Incytr(object = Data.input@assays$RNA@data,
                            meta = Data.input@meta.data,
                            sender = Sender.group,
                            receiver = Receiver.group,
                            group.by = "Type",
                            condition = c("5X", "WT"),
                            assay = NULL,
                            do.sparse = T)
    
    # Infer paths
    Incytr <- pathway_inference(Incytr,
                                DB = DB.M,
                                gene.use_Sender = DG.Sender,
                                gene.use_Receiver = DG.Receiver,
                                ligand = NULL,
                                receptor = NULL,
                                em = NULL,
                                target = NULL)
    
    # Calculate the average expression level
    Incytr <- Expr_bygroup(Incytr,
                           mean_method = NULL)
    
    # Predict signaling prob. (from transcriptomics data)
    # "cutoff_SigWeight = 0.10": keep the pathways that SigWeight>=0.10 in at least one condition
    try.function <- try(     Incytr <- Cal_SigProb(Incytr, K = 0.5, N = 2, cutoff_SigProb = 0.10,
                                                   correction = 0.001, q = NULL) )
    if( "try-error" %in% class(try.function) ){
      if( isTRUE(grepl(text, try.function[1])) ){
        next
      }
    }
    
    # Fold change of each gene between conditions (optional, does not affect the score calculation, can be commented out to save time)
    # Incytr <- Cal_scFC(Incytr, 
    #                    count.matrix = Data.input@assays$RNA@counts,
    #                    pseudocount = TRUE, 
    #                    fitType='mean')
    
    # Integrate the multiomics data (depend on the Scenario)
    if(Scenario == 1){ # no multimodel data
      Incytr <- Integr_multiomics(Incytr, 
                                  pr.data_condition1 = NULL,
                                  pr.data_condition2 = NULL,
                                  pr.correction = NULL,
                                  pr.q = NULL,
                                  ps.data_condition1 = NULL,
                                  ps.data_condition2 = NULL,
                                  ps.correction = NULL,
                                  ps.q = NULL,
                                  py.data_condition1 = NULL,
                                  py.data_condition2 = NULL,
                                  py.correction = NULL,
                                  py.q = NULL)
    }else if(Scenario == 2){ # proteomics data
      Incytr <- Integr_multiomics(Incytr, 
                                  pr.data_condition1 = pr_5X,
                                  pr.data_condition2 = pr_WT,
                                  pr.correction = NULL,
                                  pr.q = NULL,
                                  ps.data_condition1 = NULL,
                                  ps.data_condition2 = NULL,
                                  ps.correction = NULL,
                                  ps.q = NULL,
                                  py.data_condition1 = NULL,
                                  py.data_condition2 = NULL,
                                  py.correction = NULL,
                                  py.q = NULL)
    }else if(Scenario == 3){ # ps data
      Incytr <- Integr_multiomics(Incytr, 
                                  pr.data_condition1 = pr_5X,
                                  pr.data_condition2 = pr_WT,
                                  pr.correction = NULL,
                                  pr.q = NULL,
                                  ps.data_condition1 = ps_5X,
                                  ps.data_condition2 = ps_WT,
                                  ps.correction = NULL,
                                  ps.q = NULL,
                                  py.data_condition1 = NULL,
                                  py.data_condition2 = NULL,
                                  py.correction = NULL,
                                  py.q = NULL)
    }else if(Scenario == 4){# py data
      Incytr <- Integr_multiomics(Incytr, 
                                  pr.data_condition1 = pr_5X,
                                  pr.data_condition2 = pr_WT,
                                  pr.correction = NULL,
                                  pr.q = NULL,
                                  ps.data_condition1 = ps_5X,
                                  ps.data_condition2 = ps_WT,
                                  ps.correction = NULL,
                                  ps.q = NULL,
                                  py.data_condition1 = py_5X,
                                  py.data_condition2 = py_WT,
                                  py.correction = NULL,
                                  py.q = NULL)
    }
    
    # Calculate the final score
    # "cutoff_RNA_score": keep the pathways that "abs(RNA_score)>=cutoff_RNA_score"
    try.function <- try(  Incytr <- Pathway_evaluation(Incytr, score.weight = NULL, k_logi = 2, style = NULL, abs.value = NULL,
                                                       cutoff_TPRS = NULL) )
    if( "try-error" %in% class(try.function) ){
      if( isTRUE(grepl(text, try.function[1])) ){
        next
      }
    }
    
    Incytr <- Integr_kinasedata(Incytr,
                                kldata = kldata,
                                mean_method = NULL,
                                cell_group = levels(Incytr@meta$Type),
                                fold_threshold = 10)
    
    # Combine the Kinase score with previous calculated "final score"
    # "cutoff_final_score = 0.2": keep the pathways that "abs(final_score)>=0.2"
    try.function <- try( Incytr <- Cal_PRS(Incytr, KPRS.weight = 0.5,
                                           cutoff_PRS = 0.2) )
    if( "try-error" %in% class(try.function) ){
      if( isTRUE(grepl(text, try.function[1])) ){
        next
      }
    }
    
    # "cutoff_p_value = 0.1": keep the pathways that "p_value<=0.1" in at least one condition
    try.function <- try( Incytr <- Permutation_test(Incytr,
                                                    nboot = 100,
                                                    seed.use = 1L,
                                                    mean_method = NULL,
                                                    cutoff_p_value = 0.1) )
    if( "try-error" %in% class(try.function) ){
      if( isTRUE(grepl(text, try.function[1])) ){
        next
      }
    }
    
    # infer SrK (a seperate data.frame which includes both SiK and SrK, saved at Incytr@kl.explore)
    Incytr <- Kinase_exploration(Incytr,
                                 kldata = kldata,
                                 mean_method = NULL, 
                                 cell_group = levels(Incytr@meta$Type),
                                 fold_threshold = 10)
    
    # Export the results (does not include the SrK analysis results)
    output <- Export_results(Incytr,
                             indicator = TRUE)
    
    Pairwise_pathway_table[[k]] = output
    Pairwise_kinase_table[[k]] = Incytr@kl.explore
    names(Pairwise_pathway_table)[k] = paste0(Sender.group, "_to_", Receiver.group)
    names(Pairwise_kinase_table)[k] = paste0(Sender.group, "_to_", Receiver.group)
    
    k = k+1
    
    write.csv(output, file=paste0("sce4_DE_HE_", Incytr@sender, "_to_", Incytr@receiver, "_121024.csv"))
    
    if(nrow(Incytr@kl.explore)>0){
      write.csv(Incytr@kl.explore, file=paste0("sce4_DE_HE_", Incytr@sender, "_to_", Incytr@receiver, "_kinase_121024.csv"))
    }
  }
}

end.time <- Sys.time()
time.1 <- end.time - start.time
```

Check the running time
```{r}
time.1
```


Merge into one table and save
```{r}
setwd("C:/Users/chang/Documents/Research_2023/CST_project/CST_Project/output_DEG_HEG")

# "select.top = NULL": for each sender-receiver pair, save ALL pathways
dfmerge1 <- Merge_results(results.list = Pairwise_pathway_table, cutoff = NULL, group = "Both", select.top = NULL)
write.csv(dfmerge1, file=paste0("sce4_DE_HE_Allpathway_table_121024.csv"))

# "select.top = 300": for each sender-receiver pair, select the top 300 pathways with highest & lowest final scores, respectively
dfmerge2 <- Merge_results(results.list = Pairwise_pathway_table, cutoff = NULL, group = "Both", select.top = 300)
write.csv(dfmerge2, file=paste0("sce4_DE_HE_Top300_table_121024.csv"))

# save the results just in case we need to regenerate the "Pairwise_pathway_table" table
saveRDS(Pairwise_pathway_table, "sce4_DE_HE_Pairwise_pathway_table_121024.rds")
saveRDS(Pairwise_kinase_table, "sce4_DE_HE_Pairwise_kinase_table_121024.rds")
```

################################################################################

# Incytr analysis 2: input genes = DEG + HEG + proteomics + kinases

## Identify highly expressed kinases
```{r}
gene.measured = rownames(Data.input@assays$RNA@data)
kinase.measured = unique(kldata[kldata$gene %in% gene.measured & kldata$motif.geneName %in% gene.measured, ]$motif.geneName)

Data.kinase = subset(Data.input, features = kinase.measured)

Kinase_Exp = dataprepare_Expr_bygroup(object = Data.kinase@assays$RNA@data, 
                                      meta = Data.kinase@meta.data, 
                                      group.by = "Type", 
                                      conditions = c("5X", "WT"),
                                      omics = NULL,
                                      mean_method = NULL)
```



Vareables to store pairwise-analysis results
```{r}
Pairwise_pathway_table = list()
Pairwise_kinase_table = list()
```


```{r, include=FALSE}
start.time <- Sys.time()

setwd("C:/Users/chang/Documents/Research_2023/CST_project/CST_Project/output_DEG_HEG_proteomicsgene_kinase")

cell.groups = levels(Data.input$Type)

k = 1

for (i in 1:length(cell.groups) ) {
  for (j in 1:length(cell.groups) ) {
    # select the sender & receiver groups
    Sender.group = cell.groups[i]
    Receiver.group = cell.groups[j]
    # set the Scenario
    Scenario = 4
    # input genes are DEG
    DG.Sender <- unique(Allmarkers$gene[Allmarkers$cluster %in% c( paste0(Sender.group, "_5X"),paste0(Sender.group, "_WT") )])
    DG.Receiver <- unique(Allmarkers$gene[Allmarkers$cluster %in% c( paste0(Receiver.group, "_5X"),paste0(Receiver.group, "_WT") )])
    
    ############################################################################
    # Add highly expressed genes
    HG.Sender <- HEG.df$gene_symbol[HEG.df$cluster == Sender.group ]
    HG.Receiver <- HEG.df$gene_symbol[HEG.df$cluster == Receiver.group ]
    
    DG.Sender <- unique(c(DG.Sender, HG.Sender))
    DG.Receiver <- unique(c(DG.Receiver, HG.Receiver))
    
    ############################################################################
    # Add proteomics genes
    pr_gene.Sender = pr.gene_output$gene_symbol[ pr.gene_output$cluster == Sender.group ]
    pr_gene.Receiver = pr.gene_output$gene_symbol[ pr.gene_output$cluster == Receiver.group ]
    
    ############################################################################
    # check if the total number of genes is over 2000
    DG.Sender_new = unique(c(DG.Sender, pr_gene.Sender))
    if( length( DG.Sender_new )>2000 ){
      D = length( DG.Sender_new ) - 2000
      D1 = round(length(DG.Sender)/(length(DG.Sender)+length(pr_gene.Sender))*D)
      D2 = round(length(pr_gene.Sender)/(length(DG.Sender)+length(pr_gene.Sender))*D)
      DG.Sender = DG.Sender[1:(length(DG.Sender)-D)]
      pr_gene.Sender = pr_gene.Sender[1:(length(pr_gene.Sender)-D)]
      DG.Sender_new = unique(c(DG.Sender, pr_gene.Sender))
    }
    
        
    DG.Receiver_new = unique(c(DG.Receiver, pr_gene.Receiver))
    if( length( DG.Receiver_new )>2000 ){
      D = length( DG.Receiver_new ) - 2000
      D1 = round(length(DG.Receiver)/(length(DG.Receiver)+length(pr_gene.Receiver))*D)
      D2 = round(length(pr_gene.Receiver)/(length(DG.Receiver)+length(pr_gene.Receiver))*D)
      DG.Receiver = DG.Receiver[1:(length(DG.Receiver)-D)]
      pr_gene.Receiver = pr_gene.Receiver[1:(length(pr_gene.Receiver)-D)]
      DG.Receiver_new = unique(c(DG.Receiver, pr_gene.Receiver))
    }
    
    ############################################################################
    # add kinases
    kinase_gene.Sender = data.frame(gene_symbol = Kinase_Exp[[1]]$gene_symbol,
                                    condition1 = Kinase_Exp[[1]][ , Sender.group ],
                                    condition2 = Kinase_Exp[[2]][ , Sender.group ])
    kinase_gene.Sender = kinase_gene.Sender[ (kinase_gene.Sender$condition1 >0 | 
                                                kinase_gene.Sender$condition2 >0), ]$gene_symbol
    
    kinase_gene.Receiver = data.frame(gene_symbol = Kinase_Exp[[1]]$gene_symbol,
                                    condition1 = Kinase_Exp[[1]][ , Receiver.group ],
                                    condition2 = Kinase_Exp[[2]][ , Receiver.group ])
    kinase_gene.Receiver = kinase_gene.Receiver[ (kinase_gene.Receiver$condition1 >0 | 
                                                    kinase_gene.Receiver$condition2 >0), ]$gene_symbol
    
    DG.Sender_new = unique(c(DG.Sender_new, kinase_gene.Sender))
    DG.Receiver_new = unique(c(DG.Receiver_new, kinase_gene.Receiver))
    
    # Create object
    Incytr <- create_Incytr(object = Data.input@assays$RNA@data,
                            meta = Data.input@meta.data,
                            sender = Sender.group,
                            receiver = Receiver.group,
                            group.by = "Type",
                            condition = c("5X", "WT"),
                            assay = NULL,
                            do.sparse = T)
    
    # Infer paths
    Incytr <- pathway_inference(Incytr,
                                DB = DB.M,
                                gene.use_Sender = DG.Sender_new,
                                gene.use_Receiver = DG.Receiver_new,
                                ligand = NULL,
                                receptor = NULL,
                                em = NULL,
                                target = NULL)
    
    # Calculate the average expression level
    Incytr <- Expr_bygroup(Incytr,
                           mean_method = NULL)
    
    # Predict signaling prob. (from transcriptomics data)
    # "cutoff_SigWeight = 0.10": keep the pathways that SigWeight>=0.10 in at least one condition
    try.function <- try(     Incytr <- Cal_SigProb(Incytr, K = 0.5, N = 2, cutoff_SigProb = 0.10,
                                                   correction = 0.001, q = NULL) )
    if( "try-error" %in% class(try.function) ){
      if( isTRUE(grepl(text, try.function[1])) ){
        next
      }
    }
    
    # Fold change of each gene between conditions (optional, does not affect the score calculation, can be commented out to save time)
    # Incytr <- Cal_scFC(Incytr, 
    #                    count.matrix = Data.input@assays$RNA@counts,
    #                    pseudocount = TRUE, 
    #                    fitType='mean')
    
    # Integrate the multiomics data (depend on the Scenario)
    if(Scenario == 1){ # no multimodel data
      Incytr <- Integr_multiomics(Incytr, 
                                  pr.data_condition1 = NULL,
                                  pr.data_condition2 = NULL,
                                  pr.correction = NULL,
                                  pr.q = NULL,
                                  ps.data_condition1 = NULL,
                                  ps.data_condition2 = NULL,
                                  ps.correction = NULL,
                                  ps.q = NULL,
                                  py.data_condition1 = NULL,
                                  py.data_condition2 = NULL,
                                  py.correction = NULL,
                                  py.q = NULL)
    }else if(Scenario == 2){ # proteomics data
      Incytr <- Integr_multiomics(Incytr, 
                                  pr.data_condition1 = pr_5X,
                                  pr.data_condition2 = pr_WT,
                                  pr.correction = NULL,
                                  pr.q = NULL,
                                  ps.data_condition1 = NULL,
                                  ps.data_condition2 = NULL,
                                  ps.correction = NULL,
                                  ps.q = NULL,
                                  py.data_condition1 = NULL,
                                  py.data_condition2 = NULL,
                                  py.correction = NULL,
                                  py.q = NULL)
    }else if(Scenario == 3){ # ps data
      Incytr <- Integr_multiomics(Incytr, 
                                  pr.data_condition1 = pr_5X,
                                  pr.data_condition2 = pr_WT,
                                  pr.correction = NULL,
                                  pr.q = NULL,
                                  ps.data_condition1 = ps_5X,
                                  ps.data_condition2 = ps_WT,
                                  ps.correction = NULL,
                                  ps.q = NULL,
                                  py.data_condition1 = NULL,
                                  py.data_condition2 = NULL,
                                  py.correction = NULL,
                                  py.q = NULL)
    }else if(Scenario == 4){# py data
      Incytr <- Integr_multiomics(Incytr, 
                                  pr.data_condition1 = pr_5X,
                                  pr.data_condition2 = pr_WT,
                                  pr.correction = NULL,
                                  pr.q = NULL,
                                  ps.data_condition1 = ps_5X,
                                  ps.data_condition2 = ps_WT,
                                  ps.correction = NULL,
                                  ps.q = NULL,
                                  py.data_condition1 = py_5X,
                                  py.data_condition2 = py_WT,
                                  py.correction = NULL,
                                  py.q = NULL)
    }
    
    # Calculate the final score
    # "cutoff_RNA_score": keep the pathways that "abs(RNA_score)>=cutoff_RNA_score"
    try.function <- try(  Incytr <- Pathway_evaluation(Incytr, score.weight = NULL, k_logi = 2, style = NULL, abs.value = NULL,
                                                       cutoff_TPRS = NULL) )
    if( "try-error" %in% class(try.function) ){
      if( isTRUE(grepl(text, try.function[1])) ){
        next
      }
    }
    
    Incytr <- Integr_kinasedata(Incytr,
                                kldata = kldata,
                                mean_method = NULL,
                                cell_group = levels(Incytr@meta$Type),
                                fold_threshold = 10)
    
    # Combine the Kinase score with previous calculated "final score"
    # "cutoff_final_score = 0.2": keep the pathways that "abs(final_score)>=0.2"
    try.function <- try( Incytr <- Cal_PRS(Incytr, KPRS.weight = 0.5,
                                           cutoff_PRS = 0.2) )
    if( "try-error" %in% class(try.function) ){
      if( isTRUE(grepl(text, try.function[1])) ){
        next
      }
    }
    
    # "cutoff_p_value = 0.1": keep the pathways that "p_value<=0.1" in at least one condition
    try.function <- try( Incytr <- Permutation_test(Incytr,
                                                    nboot = 100,
                                                    seed.use = 1L,
                                                    mean_method = NULL,
                                                    cutoff_p_value = 0.1) )
    if( "try-error" %in% class(try.function) ){
      if( isTRUE(grepl(text, try.function[1])) ){
        next
      }
    }
    
    # infer SrK (a seperate data.frame which includes both SiK and SrK, saved at Incytr@kl.explore)
    Incytr <- Kinase_exploration(Incytr,
                                 kldata = kldata,
                                 mean_method = NULL, 
                                 cell_group = levels(Incytr@meta$Type),
                                 fold_threshold = 10)
    
    # Export the results (does not include the SrK analysis results)
    output <- Export_results(Incytr,
                             indicator = TRUE)
    
    Pairwise_pathway_table[[k]] = output
    Pairwise_kinase_table[[k]] = Incytr@kl.explore
    names(Pairwise_pathway_table)[k] = paste0(Sender.group, "_to_", Receiver.group)
    names(Pairwise_kinase_table)[k] = paste0(Sender.group, "_to_", Receiver.group)
    
    k = k+1
    
    write.csv(output, file=paste0("sce4_DE_HE_proteomics_kinase_", Incytr@sender, "_to_", Incytr@receiver, "_121024.csv"))
    
    if(nrow(Incytr@kl.explore)>0){
      write.csv(Incytr@kl.explore, 
                file=paste0("sce4_DE_HE_proteomics_kinase_", Incytr@sender, "_to_", Incytr@receiver, "_kinase_121024.csv"))
    }
  }
}

end.time <- Sys.time()
time.1 <- end.time - start.time
```

Check the running time
```{r}
time.1
```

Merge into one table and save
```{r}
setwd("C:/Users/chang/Documents/Research_2023/CST_project/CST_Project/output_DEG_HEG_proteomicsgene_kinase")

# "select.top = NULL": for each sender-receiver pair, save ALL pathways
dfmerge1 <- Merge_results(results.list = Pairwise_pathway_table, cutoff = NULL, group = "Both", select.top = NULL)
write.csv(dfmerge1, file=paste0("sce4_DE_HE_proteomics_kinase_Allpathway_table_121024.csv"))

# "select.top = 300": for each sender-receiver pair, select the top 300 pathways with highest & lowest final scores, respectively
dfmerge2 <- Merge_results(results.list = Pairwise_pathway_table, cutoff = NULL, group = "Both", select.top = 300)
write.csv(dfmerge2, file=paste0("sce4_DE_HE_proteomics_kinase_Top300_table_121024.csv"))

# save the results just in case we need to regenerate the "Pairwise_pathway_table" table
saveRDS(Pairwise_pathway_table, "sce4_DE_HE_proteomics_kinase_Pairwise_pathway_table_121024.rds")
saveRDS(Pairwise_kinase_table, "sce4_DE_HE_proteomics_kinase_Pairwise_kinase_table_121024.rds")
```
