---
title: "Incytr_Benchmarking_exFINDER"
author: "Changhan"
date: "2024-11-11"
output: html_document
---

# Part I: Load databases, input datasets

Load the required libraries
```{r message=FALSE,warning=FALSE}
library(exFINDER)
library(Seurat)
library(ggplot2)
library(patchwork)
library(RColorBrewer)
library(dplyr)
library(pheatmap)
library(ggraph)
library(forcats)
library(stringr)

library(tidyverse)
library(caret)
```

Load exFINDER database
```{r}
setwd("C:/Users/chang/Documents/Research_2021_Modeling/Project1/R Package/exFINDER/data")
# load exFINDER database
load("LR_layer1_mouse.rda")
load("RTF_layer2_mouse.rda")
load("TFT_layer3_mouse.rda")

exFINDER.M <- list()
exFINDER.M[[1]] <- LR_layer1_mouse
exFINDER.M[[2]] <- RTF_layer2_mouse
exFINDER.M[[3]] <- TFT_layer3_mouse
```

Incytr database
```{r}
setwd("C:/Users/chang/Documents/Research_2023/CST_project/CST_Project/Database")
# load the exFINDER database (can also download the updated database from the exFINDER github page)
load("DB_Layer1_mouse_filtered.rda")
load("DB_Layer2_mouse_filtered.rda")
load("DB_Layer3_mouse_filtered.rda")

DB.M <- list()
DB.M[[1]] <- DB_Layer1_mouse_filtered
DB.M[[2]] <- DB_Layer2_mouse_filtered
DB.M[[3]] <- DB_Layer3_mouse_filtered
```

RNA-seq data
```{r message=FALSE,warning=FALSE}
# path to the folder
setwd("C:/Users/chang/Documents/Research_2023/CST_project/CST_Project/CST_scRNAseq_data")
Data.input <- readRDS("Az_8clusteres.RDS") 
```

For exFINDER, we should run the analysis for each condition seperately (although the results are similar).
```{r}
Data.input.5X = subset(Data.input, subset = condition == "5X")
Data.input.WT = subset(Data.input, subset = condition == "WT")
```

Marker genes
```{r message=FALSE,warning=FALSE}
setwd("C:/Users/chang/Documents/Research_2023/CST_project/CST_Project/DG_analysis")
Allmarkers = read.csv("Allmarkers_18groups.csv")
```

################################################################################

# Part II: Set-up inputs

```{r}
cell.groups = levels(Data.input$Type)
conditions = c("5X", "WT")
db.names = c("exFINDER-DB", "Incytr-DB")
```


################################################################################
# Part III: exFINDER analysis (use both exFINDER-DB & Incytr-DB)

Prepare the inputs for exFINDER analysis
```{r}
# 5X
exp.data.5X <- Data.input.5X@assays$RNA@data
meta.data.5X <- Data.input.5X@meta.data

# WT
exp.data.WT <- Data.input.WT@assays$RNA@data
meta.data.WT <- Data.input.WT@meta.data
```

For benchmarking against exFINDER, we perform the following analysis: (1) infer the pathways using exFINDER-DB only; (2) predict the signaling probability using 5X and WT data, respectively; (3) keep the pathways with signaling probability >= 0.10 (same as what we used in the Incytr analysis).
```{r, warning=FALSE}
start.time <- Sys.time()

k = 1

exFINDER.output = list()

for (i in 1:length(cell.groups) ) {
  for (j in 1:length(cell.groups) ) {
    
    Sender.group = cell.groups[i]
    Receiver.group = cell.groups[j]
    
    # in exFINDER analysis, the target genes are DG.Receiver
    DG.Sender <- unique(Allmarkers$gene[Allmarkers$cluster %in% c( paste0(Sender.group, "_5X"),paste0(Sender.group, "_WT") )])
    DG.Receiver <- unique(Allmarkers$gene[Allmarkers$cluster %in% c( paste0(Receiver.group, "_5X"),paste0(Receiver.group, "_WT") )])
    
    ############################################################################
    # Add highly expressed genes
    HG.Sender <- HEG.df$gene_symbol[HEG.df$cluster == Sender.group]
    HG.Receiver <- HEG.df$gene_symbol[HEG.df$cluster == Receiver.group]
    
    DG.Sender <- unique(c(DG.Sender, HG.Sender))
    DG.Receiver <- unique(c(DG.Receiver, HG.Receiver))
    
    target = DG.Receiver
    
    ltGRN <- get_ltGRN(Target = target, DB = exFINDER.M)
    
    output.condition = list()
   
    for (n in 1:2) {  # analysis under different conditions
      if(n==1){
        # condition 1
        exp.data <- exp.data.5X
        meta.data <- meta.data.5X
      }else{
        # condition 2
        exp.data <- exp.data.WT
        meta.data <- meta.data.WT
      }
      
      # infer exSigNet 
      exSigNet <- get_exSigNet(Graph = ltGRN,
                               Ligands = DG.Sender, # the ligands must be potential external signals
                               Exp.Data = exp.data,
                               Meta.Data = meta.data,
                               par.cutoff = c(0, 0), 
                               AG.R = Receiver.group,  
                               AG.TF = Receiver.group)
      # calculate the expression levels
      exSigNet <- get_NodeValue(Graph = exSigNet,
                                Exp.Data = exp.data,
                                Meta.Data = meta.data,
                                AG.R = Receiver.group,
                                AG.TF = Receiver.group,
                                AG.T = Receiver.group,
                                Exp.ExData = exp.data,
                                Meta.ExData = meta.data,
                                AG.ExData = Sender.group)
      # predict signaling strengths
      exSigNet <- get_EdgeWeight(Graph = exSigNet, 
                                 Kh = 0.5)
      
      ########################################################################
      # below are converting the analysis outputs to the four components pathways
      # saperate the three layers of interactions
      x = exSigNet$Graph.Edge
      x1 = x[ x$from.Role == "Ligand", c("from", "to", "Weight")]
      x2 = x[ x$from.Role == "Receptor", c("from", "to", "Weight")]
      x3 = x[ x$from.Role == "TF", c("from", "to", "Weight")]
      
      # merge layer 1&2
      y = full_join(x1, x2, by = join_by(to == from), multiple = "all" )
      colnames(y) = c("Ligand", "Receptor", "Weight.1", "TF", "Weight.2")
      # merge layer 1&2&3
      y = full_join(y, x3, by = join_by(TF == from), multiple = "all" )
      colnames(y)[6:7] = c("Target", "Weight.3")
      
      y$Path = paste0(y$Ligand, "*", y$Receptor, "*", y$TF, "*", y$Target)
      y = y[ , c("Ligand", "Receptor", "TF", "Target", "Path", "Weight.1", "Weight.2", "Weight.3")]
      
      # filter the pathways: (1) ligands must be in the DG.Sender; (2) the rest must be in the DG.Receiver
      y = y[( y$Ligand %in% DG.Sender 
              & y$Receptor %in% DG.Receiver 
              & y$TF %in% DG.Receiver ), ]
      
      # remove pathways with duplicated genes
      boolean_dup <- apply(y[ , c("Ligand", "Receptor", "TF", "Target")], 1, duplicated)
      index = apply(boolean_dup, 2, sum)
      y$index = index
      y = y[y$index == 0, 1:8]
      
      # calculate the signaling probability (this formula is the same as Incytr)
      y$SigProb = y$Weight.1*y$Weight.2*y$Weight.3
      
      ## scale the exFINDER inferred signaling probability
      #v = y$SigProb
      #v.pre <- preProcess(as.data.frame(v), method=c("range"))
      #v.scaled <- predict(v.pre, as.data.frame(v))
      #y$SigProb.2 = v.scaled$v
      
      # keep the pathways which SigProb >= 0.10 
      y2 = y[y$SigProb>=0.10, ]
      
      output.condition[[n]] = y2
    }
    names(output.condition) = conditions

    exFINDER.output[[k]] = output.condition
    names(exFINDER.output)[k] = paste0(Sender.group, "_to_", Receiver.group)
    k = k+1
  }
}

end.time <- Sys.time()
time.1 <- end.time - start.time
```


Check the running time
```{r}
time.1
```

Save the results
```{r}
setwd("C:/Users/chang/Documents/Research_2023/CST_project/CST_Project/Incytr_Benchmarking")

# save the results
saveRDS(exFINDER.output, "DE_HE_exFINDER_121024.rds")
```


Once finished running the pairwise analysis using exFINDER, we can compare it to the following three Incytr analysis results: (1) DEG; (2) DEG + Proteomics genes; (3) DEG + Proteomics genes + HEG, and for each sender-receiver pair, the following questions should be answered:
 1. For those exFINDER-only pathways, how many of them are non-differentially expressed?
 2. For those Incytr-only pathways, how many of them are recorded in the exFINDER-DB but were not picked up by exFINDER?


