---
title: "Incytr_functions_evaluation"
author: "Changhan"
date: "2024-12-01"
output: html_document
---
################################################################################
Note: This code is unpublic and intended for use within the [Incytr/Signaling networks discovery paper] project only.
Please refrain from sharing it outside the project team or using it for any other projects without prior approval from the author.
################################################################################


#' Define Logistic function
```{r}
logi <- function(x,k){
  return( 2/(1+exp(-k*x))-1 )
}
```


#' Calculate the final score
#' 'style' should be either "log2FC" or "aFC"
#' cutoff_TPDS applies to the column "TPDS": abs(TPDS)>=cutoff_TPDS
```{r}
Pathway_evaluation <- function(object, 
                               score.weight = NULL, 
                               k_logi = NULL, 
                               style = NULL, 
                               abs.value = "None",
                               cutoff_TPDS = NULL){
  # check the input 'score.weight'
  if(length(score.weight)!=3){
    score.weight = c(0.5, 0.5, 0.5)
  }
  
  # parameter for logistic function
  if(is.null(k_logi)){
    k_logi = 2
  }
  
  # check the input 'style', should be either "log2FC" or "aFC"
  if(is.null(style)){
    style = "aFC"
  }else if( !(style %in% c("log2FC", "aFC")) ){
    style = "aFC"
  }
  
  # check the input 'abs.value'
  if(is.null(abs.value)){ abs.value = "None" }
  if( abs.value == "None"){
    case = c(0, 0, 0, 0)
  }else if( abs.value == "All" ){
    case = c(1, 1, 1, 1)
  }else if( !all(is.element(abs.value, c("Ligand", "Receptor", "EM", "Target"))) ){
    stop("Should specify the component(s) that need to use the absolute value!")
  }else if("Ligand" %in% abs.value){
    case[1] = 1
  }else if("Receptor" %in% abs.value){
    case[2] = 1
  }else if("EM" %in% abs.value){
    case[3] = 1
  }else if("Target" %in% abs.value){
    case[4] = 1
  }
  
  # part 1 -- scdata (base, s0)
  if(nrow(object@SigProb)==0){
    stop("No SigProb data available, should perform the analysis first.")
  }else{
    # calculate the score 
    df = object@SigProb[, style]
    df[is.na(df)] = 0 # 'NA'
    s0 = logi(df, k = k_logi)
  }
  
  # part 2 -- pr data (add-on, s1)
  if(nrow(object@pr_FC)==0){
    print("No protein fold change data available, skip this part.")
    s1 = 0
  }else{
    # get the data 
    df = object@pr_FC[, c(paste0("Ligand_pr_", style), 
                          paste0("Receptor_pr_", style), 
                          paste0("EM_pr_", style), 
                          paste0("Target_pr_", style))]
    df[is.na(df)] = 0 # 'NA'
    # calculate the score
    s1 = 0
    for (i in 1:4) {
      if(case[i]==0){ 
        s1 = s1 + logi(df[, i], k = k_logi)/4 
      }else{
        s1 = s1 + abs(logi(df[, i], k = k_logi))/4
      }
    }
  }
  
  # part 3 -- ps data  (add-on, s2)
  if(nrow(object@ps_FC)==0){
    print("No pS fold change datan available, skip this part.")
    s2 = 0
  }else{
    # get the data 
    df = object@ps_FC[, c(paste0("Ligand_ps_", style), 
                          paste0("Receptor_ps_", style), 
                          paste0("EM_ps_", style), 
                          paste0("Target_ps_", style))]
    df[is.na(df)] = 0 # 'NA'
    # calculate the score
    s2 = 0
    for (i in 1:4) {
      if(case[i]==0){ 
        s2 = s2 + logi(df[, i], k = k_logi)/4 
      }else{
        s2 = s2 + abs(logi(df[, i], k = k_logi))/4
      }
    }
  }
  
  # part 4 -- py data  (add-on, s3)
  if(nrow(object@py_FC)==0){
    print("No pY fold change datan available, skip this part.")
    s3 = 0
  }else{
    # get the data
    df = object@py_FC[, c(paste0("Ligand_py_", style), 
                          paste0("Receptor_py_", style), 
                          paste0("EM_py_", style), 
                          paste0("Target_py_", style))]
    df[is.na(df)] = 0 # 'NA'
    # calculate the score
    s3 = 0
    for (i in 1:4) {
      if(case[i]==0){ 
        s3 = s3 + logi(df[, i], k = k_logi)/4 
      }else{
        s3 = s3 + abs(logi(df[, i], k = k_logi))/4
      }
    }
  }
  
  # calculate the final score
  s4 = s0 + score.weight[1]*s1 + score.weight[2]*s2 + score.weight[3]*s3
  object@Evaluation = data.frame(Path = object@pathways$Path, 
                                  TPDS = s0, 
                                  PPDS = s1, 
                                  PhPDS_ps = s2, 
                                  PhPDS_py = s3, 
                                  multimodel_score = s4)
  
  # filter the pathways
  if(!is.null(cutoff_TPDS) ){
    df.update = object@Evaluation[abs(object@Evaluation$TPDS) >= cutoff_TPDS,  ]
    
    if(nrow(df.update)>0){
      Path.update = df.update$Path
      object = object_update(object = object, Path = Path.update)
    }else{
      stop("No signaling pathway inferred using current filter, the filter has been ignored.")
    }
  }
  
  return(object)
}
```




#' Calculate the indicator values 

```{r}
Generate_indicator <- function(df, cutoff = NULL){
  # get the maximum of the df
  df1 = unlist(df)
  df1 = as.numeric(df1[df1 != "NA"])
  m = max(na.omit(df1))
  
  # check cutoff input
  if(is.null(cutoff)){ cutoff = 0 }
  
  # set 'NA' value
  M = max(m+1, cutoff+1)
  df[is.na(df)] = M
  # number of up-regulated components
  up = rowSums(df>cutoff & df<M)
  # number of down-regulated components
  down = rowSums(df<cutoff)
  
  output =  data.frame(up = up, down = down)
  
  return(output)
}
```



#' Merge the tables from the pairwise analysis
#' cutoff applies to the column "PDS"
```{r}
Merge_results <- function(results.list, cutoff_PDS = NULL, group = "Both", select.top = NULL){
  
 if(is.null(group)){
   group = "Both"
 }else if( !group %in% c("Sender.group", "Receiver.group", "Both")){
   stop("The input 'group' must be 'Sender.group', 'Receiver.group', or 'Both'.")
 }
  
  if(length(results.list)==1){
    print("Only one result table found.")
    return(results.list)
  }else if(length(results.list)==0){
    stop("The input 'results.list' cannot be empty.")
  }
  
  # check the colnames
  var.names = colnames(results.list[[1]])
  if(!all(c("PDS", "Sender.group", "Receiver.group") %in% var.names)){
    stop("The tables must have columns 'PDS', 'Sender.group', and 'Receiver.group'.")
  }
  
  for (i in 2:length(results.list)) {
    if( !identical(var.names, colnames(results.list[[i]])) ){
      stop("The colnames of all tables must be the same.")
    }
  }
  
  # merge all tables
  df <- results.list[[1]]
  for (i in 2:length(results.list)) {
    df <- rbind(df, results.list[[i]])
  }
  
  # apply cutoff
  if(!is.null(cutoff_PDS)){
    df <- df[abs(df$PDS) >= cutoff_PDS, ]
  }
  
  # apply select.top & group.by
  if(group == "Both"){
    df %>%
      group_by(ID_2) -> df
  }else if(group == "Sender.group"){
    df %>%
      group_by(Sender.group) -> df
  }else{
    df %>%
      group_by(Receiver.group) -> df
  }
  
  if(!is.null(select.top)){
    df %>%
      arrange(desc(df$PDS)) %>%
      slice_head(n = select.top) -> df.top
    
    df %>%
      arrange(desc(df$PDS)) %>%
      slice_tail(n = select.top) -> df.bottom
    
    df <- rbind(df.top, df.bottom)
    df <- df[!duplicated(df), ]
  }
  
  df %>%
    ungroup() -> df
  
  return(df)
}
```


#' Export the table (with kinase analysis results)
#' cutoff_SigWeight applies to the columns "SigWeight_condition1", "SigWeight_condition2": at least one column must >= cutoffSigWeight
#' cutoff_RNA_score applies to the column "RNA_score": abs(RNA_score)>cutoff_RNA_score
#' cutoff_p_value applies to the columns "p_value_condition1", "p_value_condition2": at least one column must <= cutoff_p_value
#' cutoff_final_score applies to the column "final_score": abs(final_score)>cutoff_final_score
#' select.top: select the top n rows with the highest and lowest final_score values, respectively
```{r}
Export_results <- function(object, 
                           indicator = FALSE){
  
  # obtain the pathway table
  df <- object@pathways
  df$Sender.group = object@sender
  df$Receiver.group = object@receiver
  
  if( length(object@sc_FC)==2 ){
    # export the log2FoldChange of all genes in the sender group between two conditions (Ligand)
    df_sender <- data.frame(Gene = rownames(object@sc_FC[[1]]), log2FoldChange = object@sc_FC[[1]]$log2FoldChange)
    # export the log2FoldChange of all genes in the receiver group between two conditions (R, EM, T)
    df_receiver <- data.frame(Gene = rownames(object@sc_FC[[2]]), log2FoldChange = object@sc_FC[[2]]$log2FoldChange)
    
    df <- left_join(df, df_sender, by = c("Ligand" = "Gene"))
    df <- left_join(df, df_receiver, by = c("Receptor" = "Gene"))
    df <- left_join(df, df_receiver, by = c("EM" = "Gene"))
    df <- left_join(df, df_receiver, by = c("Target" = "Gene"))
    colnames(df)[ (ncol(df)-3):ncol(df) ] = c("Ligand_sclog2FC", "Receptor_sclog2FC", "EM_sclog2FC", "Target_sclog2FC")  
  }
  
  
  # check each part and merge the results into the table 'df'
  if(nrow(object@SigProb)>0){
    if(length(object@conditions)==2){
      df <- left_join(df, object@SigProb, join_by(Path))
    }else{
      df <- left_join(df, object@SigProb[ , 1:2], join_by(Path))
    }
  }
  
  # add the p-value columns and apply the cutoff values
  if(nrow(object@p_value)>0){
    if(length(object@conditions)==2){
      df <- left_join(df, object@p_value, join_by(Path))
    }else{
      df <- left_join(df, object@p_value[ , 1:2], join_by(Path))
    }
  }else{
    message("No p-value information found.")
  }
  
  if(length(object@conditions)==1){
    if(nrow(df)==0){
      stop("No pathway inferred under current filters.")
    }
    
    df$ID_1 <- paste0(df$Path, "_", df$Sender.group, "_", df$Receiver.group)
    df$ID_2 <- paste0(df$Sender.group, "_", df$Receiver.group)
    
    return(df)
  }
  
  ##############################################################################
  
  if(nrow(object@pr_FC)>0){
    df <- cbind(df, object@pr_FC)
  }
  
  if(nrow(object@ps_FC)>0){
    df <- cbind(df, object@ps_FC)
  }
  
  if(nrow(object@py_FC)>0){
    df <- cbind(df, object@py_FC)
  }
  
  # check if we need to add the indicators
  if(is.null(indicator)){
    indicator = FALSE
  }
  # add the indicators: scFC_up/down, prFC_up/down, psFC_up/down, pyFC_up/down
  if(isTRUE(indicator)){
    # sc part
    if( length(object@sc_FC)==2 ){
      ind_sc = Generate_indicator(df[, c("Ligand_sclog2FC", "Receptor_sclog2FC", "EM_sclog2FC", "Target_sclog2FC")], cutoff = 0)
      colnames(ind_sc) = c("sc_up", "sc_down")
      df <- cbind(df, ind_sc)
    }
    # pr part
    if(nrow(object@pr_FC)>0){
      ind_pr = Generate_indicator(object@pr_FC[, c(1,3,5,7)], cutoff = 0)
      colnames(ind_pr) = c("pr_up", "pr_down")
      df <- cbind(df, ind_pr)
    }
    # ps part
    if(nrow(object@ps_FC)>0){
      ind_ps = Generate_indicator(object@ps_FC[, c(1,3,5,7)], cutoff = 0)
      colnames(ind_ps) = c("ps_up", "ps_down")
      df <- cbind(df, ind_ps)
    }
    # py part
    if(nrow(object@py_FC)>0){
      ind_py = Generate_indicator(object@py_FC[, c(1,3,5,7)], cutoff = 0)
      colnames(ind_py) = c("py_up", "py_down")
      df <- cbind(df, ind_py)
    }
  }
  
  # add kinase analysis results
  if(nrow(object@kl.pathways)>0){
    df = inner_join(df, object@kl.pathways, by = join_by(Path))
  }else{
    print("No kinase analysis result found.")
  }
  
  # add the PDS data frame ("multimodel_score", "PDS") and calculate the difference between two scores
  df <- cbind(df, object@Evaluation[ , 2:ncol(object@Evaluation)])
  
  # if(!is.null(select.top)){
  #   df %>%
  #     arrange(desc(df$final_score)) %>%
  #     slice_head(n = select.top) -> df.top
  #   
  #   df %>%
  #     arrange(desc(df$final_score)) %>%
  #     slice_tail(n = select.top) -> df.bottom
  #   
  #   df <- rbind(df.top, df.bottom)
  #   df <- df[!duplicated(df), ]
  # }
  
  if(nrow(df)==0){
      stop("No pathway inferred under current filters.")
  }
  
  df$ID_1 <- paste0(df$Path, "_", df$Sender.group, "_", df$Receiver.group)
  df$ID_2 <- paste0(df$Sender.group, "_", df$Receiver.group)
    
  return(df)
}
```


#' Add the K-PDS to the previous calculated multimodel score
#' cutoff_final_score applies to the column "final_score": abs(final_score)>cutoff_final_score
#' @param 
```{r}
Cal_PDS <- function(object,
                    KPDS.weight = NULL,
                    cutoff_PDS = NULL){
  
  # check the input 'KPDS.weight'
  if(is.null(KPDS.weight)){
    KPDS.weight = 0.5
  }
  
  if( length(object@EI)==0 ){
    object@Evaluation$PDS = object@Evaluation$multimodel_score
  }else{
    # Combine the Kinase score with previous calculated "final score"
    df = data.frame(score03 = object@Evaluation[ , "multimodel_score"],
                    s4_1 = object@kl.pathways[, paste0("SiK_score_", object@conditions[1])],
                    s4_2 = object@kl.pathways[, paste0("SiK_score_", object@conditions[2])]) 
    df$score04 = df$score03
    
    df$score04[df$score03>0] = df$score04[df$score03>0] + KPDS.weight*df$s4_1[df$score03>0]
    df$score04[df$score03==0] = df$score04[df$score03==0] + KPDS.weight*(df$s4_1[df$score03==0]-df$s4_2[df$score03==0])
    df$score04[df$score03<0] = df$score04[df$score03<0] - KPDS.weight*df$s4_2[df$score03<0]
    
    object@Evaluation$PDS = df$score04
  }
  
  # filter the pathways
  if(!is.null(cutoff_PDS) ){
    df.update = object@Evaluation[ abs(object@Evaluation$PDS) >= cutoff_PDS, ]
    
    if(nrow(df.update)>0){
      Path.update = df.update$Path
      object = object_update(object = object, Path = Path.update)
    }else{
      stop("No signaling pathway inferred using current filter, the filter has been ignored.")
    }
  }
  
  return(object)

}
```



