---
title: "Incytr_functions_object"
author: "Changhan"
date: "2024-02-04"
output: html_document
---

################################################################################
Note: This code is unpublic and intended for use within the [Incytr/Signaling networks discovery paper] project only.
Please refrain from sharing it outside the project team or using it for any other projects without prior approval from the author.
################################################################################


```{r, include=FALSE}
library(Matrix)
library(data.table)
library(dplyr)
library(forcats)
library(readr)
library(Seurat)
library(stringr)
library(limma)
library(stats)
# library(boot)
```




#' The object Class
#'# Class definitions
#' @importFrom methods setClassUnion
#' @importClassesFrom Matrix dgCMatrix
#' @importClassesFrom data.table data.table
```{r}
setClassUnion(name = 'AnyMatrix', members = c("matrix", "dgCMatrix"))
setClassUnion(name = 'AnyFactor', members = c("factor", "list"))
setClassUnion(name = 'AnyDF', members = c("data.frame", "data.table"))
```

#' The key slots used in the object are described below; adapted from CellChat https://github.com/sqjin/CellChat
#'
#' @slot data.raw raw count data matrix
#' @slot data normalized data matrix for CellChat analysis (Genes should be in rows and cells in columns)

```{r}
object <- methods::setClass("Incytr",
                              slots = c(data.raw = 'AnyMatrix',
                                        data = 'AnyMatrix',
                                        expr.bygroup = 'list',
                                        pathways = "data.frame",
                                        pathways_5steps = "data.frame",
                                        meta = "data.frame",
                                        sender = 'character',
                                        receiver = 'character',
                                        idents = "AnyFactor",
                                        conditions = 'character',
                                        SigProb = "data.frame",
                                        p_value = "data.frame",
                                        sc_FC = 'list',
                                        pr_FC = "data.frame",
                                        ps_FC = "data.frame",
                                        py_FC = "data.frame",
                                        Evaluation = "data.frame",
                                        kl = "data.frame",
                                        kl.pathways = "data.frame",
                                        EI = 'list',
                                        kl.explore = "data.frame",
                                        options = "list")
)
```


#' Create a new X object from a data matrix
#' Create a new X object from a data matrix; adapted from CellChat https://github.com/sqjin/CellChat
#' @param object a normalized (NOT count) data matrix (genes-by-cells)
#' @param condition a factor, which should also be included in the meta data

```{r}
create_Incytr <- function(object, 
                           meta = NULL, 
                           sender = NULL,
                           receiver = NULL,
                           group.by = NULL, 
                           conditions = NULL,
                           assay = NULL, 
                           do.sparse = T){
  
  # data matrix as input
  if (inherits(x = object, what = c("matrix", "Matrix", "dgCMatrix"))) {
    message("Create an Incytr object from a data matrix")
    data <- object
    if (is.null(group.by)) {
      group.by <- "labels"
    }
  }
  
  if ( is.null(sender) | is.null(receiver) ){
    stop("Please specify the 'sender' and 'receiver'!")
  }else if ( !(sender %in% levels(factor(meta[, group.by]))) | !(receiver %in% levels(factor(meta[, group.by]))) ){
    stop("The inputs 'sender' and 'receiver' must be the clusters!")
  }
  
  if (!is.null(meta)) {
  if (inherits(x = meta, what = c("matrix", "Matrix"))) {
       meta <- as.data.frame(x = meta)
  }
  if (!is.data.frame(meta)) {
    stop("The input `meta` should be a data frame")
  }
  if (!identical(rownames(meta), colnames(data))) {
      cat("The cell barcodes in 'meta' is ", head(rownames(meta)),'\n')
      warning("The cell barcodes in 'meta' is different from those in the used data matrix.
              We now simply assign the colnames in the data matrix to the rownames of 'mata'!")
      rownames(meta) <- colnames(data)
    }
  } else {
    meta <- data.frame()
  }
  
  if( !"condition" %in% colnames(meta) ){
    meta$condition = "condition1"
  }
  
  meta$condition <- factor(meta$condition)
  if (is.null(conditions)){
    conditions <- as.character(levels(meta$condition))
  }
  
  idents = factor(meta[, group.by])
  names(idents) <- colnames(data)
  
   object <- methods::new(Class = "Incytr",
                         data = data,
                         meta = meta,
                         sender = sender,
                         receiver = receiver,
                         idents = idents,
                         conditions = conditions)
  object@options$mode <- "single"
  return(object)
  
}
```

